How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-15 tse-paper/content.tex
correct first half of lit survey

% TODO: move down to where mentioned in detail
% \begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
% {\textbf{RQ1:} What is the state-of-research on build
% log analysis?}
% \begin{itemize}
%   \item[\textbf{RQ1.1:}] Which information is targeted?
%   \item[\textbf{RQ1.2:}] Which techniques are used?
%   \item[\textbf{RQ1.3:}] Which build logs are analyzed?
% \end{itemize}
% \end{simplebox}

@@

\begin{figure}[htb]
	\centering
	\includegraphics[width=\columnwidth, clip]{img/lit_survey.pdf}
	\caption{Our literature selection process, following
	\cite{petersen2015guidelines}.}
	\label{fig:lit-survey}
\end{figure}

The field of analyzing build
logs is relatively young; We have seen scattered
research efforts on build log analysis.
However,the processing of build logs was often
a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs.
We conducted a systematic mapping study to determine the
state-of-research on build log analysis following the guidelines
by Petersen et al.~\cite{petersen2015guidelines}.

In the following section, we explain how build log analysis differs
from system log analysis.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  keywordstyle=\color{Plum},
  escapeinside=**
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet Responder: Received
block blk_3587508140 of size 67108 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140
terminating
  \end{lstlisting}
  \caption{System Log excerpt.
Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

The techniques developed for system log analysis can also be applied
to build logs.
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
In this article, we focus on extracting a single specified information
from the build log as a whole with chunk retrieval techniques.
Chunk
retrieval techniques are used as a part of log parsing to retrieve the
values of variable parts in a log message, e.g.\ by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\subsection{Study Selection}
% TODO shorten
\Cref{fig:lit-survey} presents an overview of our study selection.
In this first literature overview study on analyzing build logs, we
 survey as broad a population of scientific material as possible.
We
 do not limit our sources to specific Software Engineering venues, as
 is often done~\cite{petersen2015guidelines}, but survey all sources
 monitored by Google Scholar, including Bachelor's and Master's
 theses, academic slide decks, and books, since the field of build log
 analysis is so young.
We use Google Scholar as opposed to Scopus or
 other databases, as it is the search engine for acdemic articles with
 the broadest and most current index, including preprints from sources
 such as arXiv.
% TODO: maybe remove, this screams snowballing needed
One challenge is that no common
nomenclature has been established yet (in fact, this is one of the
goals of this article).

% TODO Caro cite guidelines
Our Google Scholar
search for variations ``build log'' returns more than 2.2 million results
(2020-03-15).
We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works.
We explicitly decided to exclude systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 combined search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

@@

First, we filtered articles based on (1) their title and (2) the
contextual information Google Scholar displayed.
We inlcuded only
publications written in a language intelligible to the authors (that
is, English or German), which disregarded fewer than 1\% of search
hits.
Our strategy for this round was to err on the side of
caution, that is rather including a paper than not.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

With these 256 remaining articles, we followed a backward snowball
sampling for related work.
This meant that we added an article if referred in them in the context
of log analysis.

This added five papers (eight before
duplicate elimination) to our literature set, coming to a total of 261
articles.

Subsequently, we did a second, finer filter phase by downloading the
full articles, reading their abstracts, and searching the full text
for the occurrence of ``log''.
If these papers showed traces of
working with build logs, we included them in our survey.
In total, we
were left with 87 articles after this search.

\subsection{Article Classification}
Having trimmed down the set of articles, we investigated them in depth.
We kept all articles that use logs from software builds as input to
some kind of process and excluded another 15 articles.
For several cases, multiple articles were extensions of others.
Of these we only considered the earliest article which reported on
the processing of build logs.
This excluded 11 articles from our survey.
In sum, we further classified 61 articles.

We wanted to characterize how researchers use buildlogs as input to
any processes.
We were interesed in
\begin{itemize}
  \item what kind of build logs they are using and the source of these
  logs.
  \item what information they are retrieving from the logs,
  its kind (chunk/classification) and
  what they use the retrieved information for
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their
  tools
\end{itemize}
To objecifiy our data extraction we created a template of the targeted
factors and performed a pilot.
Two authors extracted the data from five articles and we held a
consensus meeting to unify our understanding
of the data extraction template.
The remaining articles were divided over these two authors.
We inspected the articles as closely as necessary,
starting from occurences of ``log'' or ``build'' within the text.

The full data extraction template and our results of this step can be
found in our replication package~\cite{brandt2020chunk-replication}.
Note that for most questions we allowed assigning multiple values,
e.g.\ if
two kinds of informations were targeted during a build log analysis.

\subsection{Results}
In the following, we will present the results of our literature survey.
Starting with the information that was targeted and what it was employed
for, over the techniques used until the origin of the build logs
which were analyzed.

\begin{table*}[tbhp]
\centering
\caption{Overview of buildlog analysis techniques.}
\begin{tabularx}{\textwidth}{@{}lXl@{}}
\toprule
Name			     & Sources	& Frequency	  \\
\midrule
Parser	&
~{\smallish
\cite{vassallo2018un-break,zhang2016android,seo2014programmers,hassan2019tackling,hassan2017automatic,chromy2007integration,mesbah2019deepdelta,wen2018blimp,kwon2018prioritizing,adams2007design,rahman2018impact,brandyberry2006continuous,tomassi2019bugswarm,ren2018automated,vassallo2019automated,cavalcanti2019impact,sippola2013qt,felipe2012towards,shi2018evaluating,urli2018design,selberg2012use}}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_21.pdf}
\\
Regular expression
&\cite{beller2017oops,hassan2017change,macho2018automatically,vassallo2017a-tale,lou2019history,hassan2017automatic,rott2019empirische,zampetti2019study,zhao2018comparing,rausch2017empirical,ghaleb2019studying,zampetti2017open,zhang2019large,kavaler2019tool,morris2010experience}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_15.pdf}
\\
Manual inspection  &
\cite{sulir2016quantitative,hassan2017automatic,bouabana2019theory,barinov2017applying,silva2018build,ghaleb2019empirical,marcozzi2019systematic,hukkanen2015adopting,rausch2017empirical,hassan2017mining,zolfagharinia2017not,cassee2019impact}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_12.pdf}
\\
Machine Learning  &
\cite{hassan2017change,lou2019history,lindqvist2019detection,ren2018automated,schulz2017active}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_5.pdf}
\\
\makecell[cl]{{\smallish Natural Language Processing}}	&
\cite{hassan2017change,lou2019history,schulz2017active} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Information Retrieval  &
\cite{hassan2017change,lindqvist2019detection,ren2018automated} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Analysis  &
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Keyword Search	&
\cite{brandyberry2006continuous,zhang2019large,kavaler2019tool} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Scan  & \cite{clemencic2014new,hibbard2001visualization} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_2.pdf}
\\
Other  &
\cite{zhang2016android,hassan2017change,lou2019history,silva2018build,ren2018automated,schulz2017active}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_10.pdf}
\\
None identified  &
\cite{macho2017preventing,felipe2012towards,orellana2017differences,madeyski2017continuous,zhao2017impact,santolucito2018statically,makihara2018multi,mcintosh2012evolution,gallaba2018noise,matthies2016scrumlint}
&
\\
\bottomrule
\end{tabularx}
\label{tab:litsur:techniques}
\end{table*}

@@

@@

% TODO update RQs
% TODO incorporate these RQs into the text where they are being answered

This section presents the answers to our research questions:
\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which build log analysis techniques exist?
  \item[\textbf{RQ2:}] Which criteria influence the suitability of a
  chunk retrieval technique for build logs?

  {\small \item[\textbf{~~ RQ2.1:}] How many examples do PBE, CTS,
  and KWS need to perform best?
  \item[\textbf{~~ RQ2.2:}] How structurally similar do the examples
  for PBE, CTS and KWS need to be for the techniques to be applicable?
  \item[\textbf{~~ RQ2.3:}] How accurate are the retrievals of PBE, CTS,
  and KWS?}
\end{itemize}
\end{simplebox}

>>>>

% also moved down to where mentioned in detail
% left copy here for convenience

% \begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
% {\textbf{RQ1:} What is the state-of-research on build
% log analysis?}
% \begin{itemize}
%   \item[\textbf{RQ1.1:}] Which information is targeted?
%   \item[\textbf{RQ1.2:}] Which techniques are used?
%   \item[\textbf{RQ1.3:}] Which build logs are analyzed?
% \end{itemize}
% \end{simplebox}

@@

\begin{figure}[hb]
	\centering
	\includegraphics[width=\columnwidth, clip]{img/lit_survey.pdf}
	\caption{Our literature selection process, following
	\cite{petersen2015guidelines}.}
	\label{fig:lit-survey}
\end{figure}

The field of analyzing build
logs is relatively young; We have seen scattered
research efforts on build log analysis.
However,the processing of build logs was often
a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs.
We conducted a systematic mapping study to determine the
state-of-research on build log analysis following the guidelines
by Petersen et al.~\cite{petersen2008systematic,petersen2015guidelines}.

In the following section, we explain how build log analysis differs
from system log analysis.
Further, we describe how we selected the articles for our literature
survey and the data extraction from the resulting articles.
We present our results on how buildlogs are analyzed in literature
and discuss the implecations of these findings on our research.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  keywordstyle=\color{Plum},
  escapeinside=**
}
\begin{figure*}[!tb]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet Responder: Received
block blk_3587508140 of size 67108 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140
terminating
  \end{lstlisting}
  \caption{System Log excerpt.
Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

The techniques developed for system log analysis can also be applied
to build logs.
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
However, many of the techniques developed to analyze system logs
leverage their inherent event structure.
As build logs lack this sturcture, we focussed our literature
survey on articles that as a source of information.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing to retrieve the
% values of variable parts in a log message, e.g.\ by using regular
% expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of our study selection
process.
Since the field of build log analysis is so young, we
surveyed a population of scientific material as broad as possible.
We did not limit our sources to specific Software Engineering venues, as
is often done~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's
theses and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases, as it is the search engine for acdemic articles with
the broadest and most current index, including preprints.

Our Google Scholar search for variations of ``build log''
returns more than 2.2 million results (2020-03-15).
We refined our search criterions to exclude obvious
non-related works.
For example, we decided to exclude systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

@@

First, we filtered articles based on (1) their title and (2) the
contextual information Google Scholar displayed.
We inlcuded only
publications written in a language intelligible to the authors (that
is, English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

With these 256 remaining articles, we followed a backward snowball
sampling for related work.
If one one of the articles referenced a new article in the context
of build log analysis we added the new to our literature set.
This added five papers (eight before
duplicate elimination), coming to a total of 261
articles.

Subsequently, we performed a second, finer filter phase by downloading the
full articles, reading their abstracts, and searching the full text
for the occurrence of ``log''.
If these papers showed traces of
working with build logs, we included them in our survey.
In total, we
were left with 87 articles after the two filtering phases.

\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-research on build
log analysis?}
\begin{itemize}
  \item[\textbf{RQ1.1:}] Which information is targeted?
  \item[\textbf{RQ1.2:}] Which techniques are used?
  \item[\textbf{RQ1.3:}] Which build logs are analyzed?
\end{itemize}
\end{simplebox}

\subsection{Article Classification}
Having trimmed down the set of articles, we investigated them in depth.

We wanted to characterize how researchers use buildlogs as input to
any processes.
We were interesed in
\begin{itemize}
  \item what information they are retrieving from the logs and
  what they use the retrieved information for (\textbf{RQ1.1})
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their
  tools (\textbf{RQ1.2})
  \item what kind of build logs they are using and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}
To objecifiy our data extraction we created a template of the targeted
factors and performed a pilot, as reccommended by Petersen et
al.~\cite{petersen2015guidelines}.
Two authors extracted the data from five articles and we held a
consensus meeting to unify our understanding
of the data extraction template.
The remaining articles were divided over these two authors.
We inspected the articles as closely as necessary,
starting from occurences of ``log'' or ``build'' within the text.

Relevant for our survey were all articles that
use logs from software builds as input to
some kind of process.
This excluded 15 articles from our study.
For several cases, multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the processing of build logs.
This excluded another 11 articles.
In sum, we further extracted data from 61 articles.

The full data extraction template and our results of this step can be
found in our replication package~\cite{brandt2020chunk-replication}.
Note that for most questions we allowed assigning multiple values,
e.g.\ if
two kinds of information were targeted during a build log analysis.

\subsection{Results}
In the following, we will present the results of our literature survey.
We present the information within the build logs that was targeted,
the retrieval techniques used and the origin of the build logs
which were analyzed.

\begin{table*}[tbhp]
\centering
\caption{Overview of build log analysis techniques.}
\begin{tabularx}{\textwidth}{@{}lXl@{}}
\toprule
Name			     & Sources	& Frequency	  \\
\midrule
Parser	&
~{\smallish
\cite{vassallo2018un-break,zhang2016android,seo2014programmers,hassan2019tackling,hassan2017automatic,chromy2007integration,mesbah2019deepdelta,wen2018blimp,kwon2018prioritizing,adams2007design,rahman2018impact,brandyberry2006continuous,tomassi2019bugswarm,ren2018automated,vassallo2019automated,cavalcanti2019impact,sippola2013qt,felipe2012towards,shi2018evaluating,urli2018design,selberg2012use}}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_21.pdf}
\\
Regular expression
&\cite{beller2017oops,hassan2017change,macho2018automatically,vassallo2017a-tale,lou2019history,hassan2017automatic,rott2019empirische,zampetti2019study,zhao2018comparing,rausch2017empirical,ghaleb2019studying,zampetti2017open,zhang2019large,kavaler2019tool,morris2010experience}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_15.pdf}
\\
Manual inspection  &
\cite{sulir2016quantitative,hassan2017automatic,bouabana2019theory,barinov2017applying,silva2018build,ghaleb2019empirical,marcozzi2019systematic,hukkanen2015adopting,rausch2017empirical,hassan2017mining,zolfagharinia2017not,cassee2019impact}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_12.pdf}
\\
Machine Learning  &
\cite{hassan2017change,lou2019history,lindqvist2019detection,ren2018automated,schulz2017active}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_5.pdf}
\\
\makecell[cl]{{\smallish Natural Language Processing}}	&
\cite{hassan2017change,lou2019history,schulz2017active} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Information Retrieval  &
\cite{hassan2017change,lindqvist2019detection,ren2018automated} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Analysis  &
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Keyword Search	&
\cite{brandyberry2006continuous,zhang2019large,kavaler2019tool} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
Scan  & \cite{clemencic2014new,hibbard2001visualization} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_2.pdf}
\\
Other  &
\cite{zhang2016android,hassan2017change,lou2019history,silva2018build,ren2018automated,schulz2017active}
&
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_10.pdf}
\\
None identified  &
\cite{macho2017preventing,felipe2012towards,orellana2017differences,madeyski2017continuous,zhao2017impact,santolucito2018statically,makihara2018multi,mcintosh2012evolution,gallaba2018noise,matthies2016scrumlint}
&
\\
\bottomrule
\end{tabularx}
\label{tab:litsur:techniques}
\end{table*}

@@

\begin{simplebox}[minipage boxed title*=-2.5cm,
attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ2:} When are PBE{,} CTS or KWS suited to
retrieve chunks from build logs?}
\begin{itemize}
  \item[\textbf{RQ2.1:}] How many examples does a technique need to
  perform best?
  \item[\textbf{RQ2.2:}] How structurally similar do the examples
  need for a technique to be applicable?
  \item[\textbf{RQ2.3:}] How accurate are the retrievals of a technique?
\end{itemize}
\end{simplebox}

@@
